---
title: "GGally - ggpairs function"
author: "Collin O'Connor"
date: "9/10/2020"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(GGally)
```

## Author Background

- Bullet 1
- Bullet 2
- Bullet 3

## Uses
* Useful for univariate and bivariate analysis

* ggpairs does three things in one line of code
  * Scatterplot specific variables to examine for correlation

  * Creates a density graph for single variables

  * Tests for correlation

## Univariate analysis
* Needed to determine distribution of variables

* Allows researcher to do proper analysis

  * Basic 1st step when looking at a variable is a histogram
  
```{r hist, echo=TRUE}
hist(mtcars$mpg, main="mtcars MPG distribution",xlab="Miles per Gallon")
#Shows this variable is slightly skewed, but probably normal enough
```


## Bivariate analysis
* May want to test for correlation 
  * Base R can do a test of correlation
  
```{r corr, echo=TRUE}
cor.test(mtcars$mpg,mtcars$disp)
```
* Also useful to plot both variables to look for trends
```{r singlescatter, echo=TRUE}
plot(mtcars$mpg,mtcars$disp,xlab="Miles per Gallon",ylab="Displacement (cu.in.)",main="Relating MPG to Displacement in mtcars dataset")
```

## Trying to plot full dataset using base R
```{r fullplot, echo=TRUE}
plot(mtcars[-c(2,8:11)])
```

* Returns a plot comparing all variables in a matrix

* Flips axis for each plot
  * No correlation test
  * No univariate analysis


## Using the ggpairs function
* ggpairs in GGally package combines 3 needs

  * Creates the same scatter plots as base R without flipping y axis
  
  * Shows a distribution graph of single variable
  
  * Performs a test of correlation between intersecting variables in matrix
```{r GGally, echo = TRUE}
ggpairs(mtcars[-c(2,8:11)])
# mtcars[-c(2,8:11)] are the quant. continuous variables in mtcars
```


